A Pilot Multicenter Cluster Randomized Trial to Compare The Effect of Trauma Life Support Training Programs on Patient and Provider Outcomes

### 2023-12-30 pilot-publication/lab-book.Rmd
Create outcome tables and save as html and pdf

```{r setup}
library(pbapply)
library(openssl)
library(dotenv)
library(DataExplorer)
```

>>>>

```{r setup}
library(pbapply)
library(openssl)
library(dotenv)
library(DataExplorer)
noacsr::source_all_functions()
```

## Import saved outcome results and create outcome tables

```{r}
## Load outcome results
outcome.results <- readRDS("outcome-results.rds")

## Extract cell references
cell.references <- names(outcome.results$point.estimates)

table.names <- strsplit(cell.references, split = "\\|\\|\\.") %>%
    lapply(function(table.name) table.name[1]) %>%
    unique()

ci.levels <- outcome.results$confidence.intervals[[2]][, 1]

## Combine point estimates and confidence intervals into a single table
combined.outcome.results <- lapply(ci.levels, function(ci.level) {
    lapply(cell.references, function(cell.reference) {
        point.estimate <- outcome.results$point.estimates[[cell.reference]]
        ci.data <- outcome.results$confidence.intervals[[cell.reference]]
        ci <- ci.data[ci.data[, 1] == ci.level, c("lower", "upper")]
        combined <- NA
        if (!all(is.na(ci))) {
            ## combined <- sprintf("%0.2f", c(point.estimate, ci))
            combined <- paste0(combined[1], " (", combined[2], ", ", combined[3], ")")
        }
        return(combined)
    }) %>%
        setNames(cell.references)
}) %>%
    setNames(paste0("ci.level.", ci.levels))

## Create outcome tables
create_ci_level_tables <- function(table.name, ci.data) {
    ## Extract table data
    table.data <- ci.data[grep(table.name, names(ci.data), fixed = TRUE)]
    column.names <- strsplit(names(table.data), split = "\\|\\|\\.") %>%
        lapply(function(column) column[2]) %>%
        unique() %>%
        stringr::str_replace_all(pattern = "\\|\\|column\\:\\:", replacement = "") %>%
        stringr::str_replace_all(pattern = "\\*\\*", replacement = "")
    row.names <- strsplit(names(table.data), split = "\\|\\|\\.") %>%
        lapply(function(row) row[3]) %>%
        unique() %>%
        stringr::str_replace_all(pattern = "\\|\\|row\\:\\:index\\:[0-9]*\\:\\:", replacement = "") %>%
        stringr::str_replace_all(pattern = "\\|\\|", replacement = "")

    ## Replace outcome names with labels
    all.outcomes <- all_outcomes()
    outcomes.column.list <- lapply(row.names, function(row.name) {
        new.row.name <- row.name
        index <- FALSE
        if (!is.null(all.outcomes[[row.name]])) {
            new.row.name <- all.outcomes[[row.name]]$label
            index <- TRUE
        }
        return(list(row.name = new.row.name, index = index))
    })
    outcomes.column.index <- sapply(outcomes.column.list, function(outcomes.column) outcomes.column$index)
    outcomes.column <- sapply(outcomes.column.list, function(outcomes.column) outcomes.column$row.name)

    ## Create table
    ci.table <- matrix(table.data, ncol = length(column.names)) %>%
        tibble::as_tibble(.name_repair = "minimal") %>%
        tibble::add_column(outcomes.column, .before = 1, .name_repair = "minimal") %>%
        setNames(c("Outcome", column.names))

    ## Replace NA with empty string
    ci.table[] <- lapply(ci.table, function(column) {
        column[is.na(column)] <- ""
        return(column)
    })

    ## Convert table to gt
    ci.table <- ci.table %>%
        gt::gt() %>%
        gt::tab_style(
            style = gt::cell_text(weight = "bold"),
            locations = gt::cells_body(
                columns = c(Outcome),
                rows = outcomes.column.index
            )
        ) %>%
        gt::tab_style(
            style = gt::cell_text(weight = "bold"),
            locations = gt::cells_column_labels()
        )

    ## Return table
    return(ci.table)
}

tables <- lapply(
    combined.outcome.results, function(ci.data) {
        {
            lapply(table.names, create_ci_level_tables, ci.data = ci.data)
        } %>%
            setNames(stringr::str_replace(table.names, pattern = "\\|\\|table\\:\\:", replacement = ""))
    }
)

## Save tables to file
document.string <- lapply(
    names(tables), function(ci.level.tables.name) {
        ## Extract confidence level as percentage
        ci.level <- stringr::str_remove(ci.level.tables.name, pattern = "ci\\.level\\.") %>%
            as.numeric() %>%
            scales::percent()
        ci.level.tables <- tables[[ci.level.tables.name]]
        paste0(
            "# ", ci.level, " confidence level\n\n",
            lapply(names(ci.level.tables), function(ci.level.table.name) {
                ci.level.table <- ci.level.tables[[ci.level.table.name]]
                paste0(
                    "## ", ci.level.table.name, "\n\n",
                    "\`\`\`{r, echo = FALSE} \n\n",
                    "tables[[\"", ci.level.tables.name, "\"]][[\"", ci.level.table.name, "\"]]", "\n\n",
                    "\`\`\`\n\n"
                )
            }) %>%
                paste0(collapse = "\n\n")
        )
    }
) %>%
    unlist() %>%
    paste0(collapse = "\n\n")

## Write document string to file
fileConn <- file("outcome-tables.Rmd")
writeLines(document.string, fileConn)
close(fileConn)

## Create html document
rmarkdown::render("outcome-tables.Rmd", output_format = rmarkdown::html_document(toc = TRUE))
```

The html document can then be compiled as pdf using pandoc from the command line:

```{bash}  
pandoc outcome-tables.html -o outcome-tables.pdf --pdf-engine=weasyprint
```
